Towards semantic-rich word embeddings

### 2019-03-20 paper/main.tex
Improve introduction section

Word Sense Disambiguation (WSD) is an open problem of natural language processing (NLP) and ontology. WSD is identifying which sense of a word (i.e. meaning) is used in a sentence based on the word context. Difficulty is when the word has multiple meanings (e.g. a decision tree, a tree data structure, a tree in a forest). The problem requires two inputs: a dictionary to specify the senses which are to be disambiguated and a corpus of language data to be disambiguated. WSD task has two variants: "lexical sample" and "all words" task. The former aim to disambiguate the occurrences of a small sample of selected target words, while in the latter all the words in a piece of running text need to be disambiguated. Our solution targets the former one, but could be extended to the latter variant.
The solution to WSD would be useful in many NLP related problems as: relevance of search engines, anaphora resolution, coherence, inference, etc.

Word embeddings are a product of feature learning techniques in NLP, where words from the vocabulary are mapped to vectors of real numbers. Conceptually it involves a dimensionality reduction from a space with one dimension per word to a continuous vector space with a much lower dimension.
Methods to generate this mapping include artificial neural networks\cite{word2vec}\cite{GloVe}\cite{fastText}, dimensionality reduction on the word co-occurrence matrix\cite{Levy2014NWE} and probabilistic models\cite{Globerson2007EEC}.
Word embeddings are commonly used as the input representation. They have been shown to boost the performance in NLP tasks such as syntactic parsing\cite{parsingGrammars} and sentiment analysis\cite{sentimentAnalysis}.

Word embeddings cannot distinguish between different meanings of ambiguous words by themselves. By definition, there is only one embedding for each word e.g. for word ``tree'' there is a single real-valued vector. What can be done, is to try to distinguish the meaning based on the context, in which the word was used. Then, we treat each meaning as a separate keyword, which has its own embedding.
We propose a simple method to infer the word meaning: an average of the context and the word embeddings and number of improvements to this approach in the chapter ``Our method''. Experiments with our solution are presented in the chapter ``Experiments''. To conduct those experiments we have created the dataset composed of 6 ambiguous words, with 4 to 7 meanings each, and collected real-world usage examples of those meanings, with tagged words to be disambiguated. We describe our dataset in the chapter ``Dataset''. In the chapter ``Related work'' we present other approaches to WSD.

@@

\section{Experiments}
The goal of following experiments is to check how semantic-rich keyword embeddings (\ref{sec:kw_embed}, \ref{sec:ctx_embed}) perform for the dataset we collected (\ref{sec:dataset}), both for the basic approach (\ref{sec:baseline_model}) and after applying optimization techniques (\ref{sec:optimization}).
We use a pretrained embedding model from spaCy - \textit{en\_vectors\_web\_lg}, which contains 300-dimensional word vectors trained on Common Crawl with GloVe\cite{spacymodel}.

>>>>

Word embedding methods, that map the vocabulary words to low-dimensional continuous space, have been widely applied to various natural language processing (NLP) problems.
They are commonly used as the input representation of words, replacing high-dimensional one-hot encodings, and have been shown to improve the performance in tasks such as syntactic parsing\cite{parsingGrammars} and sentiment analysis\cite{sentimentAnalysis}.

In word embedding methods, such as word2vec\cite{word2vec} or GloVe\cite{GloVe}, each word in the vocabulary has exactly one representation.
While it is enough for most words, it is problematic for ambiguous words, which can contain more than one meaning.
For example, consider the following examples with the word \textit{tree}, extracted from Wikipedia articles:

\medskip
(a) \textit{The olive, known by the botanical name Olea europaea, meaning "European olive", is a species of small \textbf{tree} in the family Oleaceae [...]}

(b) \textit{Many theories of syntax and grammar illustrate sentence structure using phrase \textbf{trees}, which provide schematics of how the words in a sentence are grouped and relate to each other.}

(c) \textit{Upon completion of listing all files and directories found, \textbf{tree} returns the total number of files and directories listed.}
\medskip

All three sentences mention the word \textit{tree} (or \textit{trees}), but the meaning differs based on context - (a) means tree as a forest plant, (b) tree as a parse tree, (c) tree as a command in Unix systems.

For many applications, such as improving relevance of search engines, anaphora resolution or coherence, identifying which meaning is used, based on context, is important.
This task is called Word Sense Disambiguation (WSD) and is an open problem in NLP domain.
Word embeddings cannot be applied to WSD out-of-the-box, since they cannot differentiate between multiple meanings of an ambiguous word.

In this work, we suggest to create semantically rich embeddings for each meaning of a word (\textit{keyword}) (\ref{sec:kw_embed}), by averaging embeddings of the ambiguous word and words describing its meaning (\ref{sec:baseline_model}).
We evaluate this approach on a small WSD task, gathered from Wikipedia articles (\ref{sec:dataset}).
Word is disambiguated by choosing the closest keyword embedding given the embedding of context, which is an average of word embeddings surrounding the word (\ref{sec:ctx_embed}).
We then propose an optimization method, to improve the quality of embeddings and the result on WSD task as well -  we move keyword embeddings closer to embeddings of contexts they appear in, using examples from the training set (\ref{sec:optimization}).
A number of experiments is run to test the proposed method and its optimized version (\ref{sec:experiments}).
Finally, we discuss our findings and present future work (\ref{sec:conclusion}).

@@

\section{Experiments}
\label{sec:experiments}
The goal of following experiments is to check how semantic-rich keyword embeddings (\ref{sec:kw_embed}, \ref{sec:ctx_embed}) perform for the dataset we collected (\ref{sec:dataset}), both for the basic approach (\ref{sec:baseline_model}) and after applying optimization techniques (\ref{sec:optimization}).
We use a pretrained embedding model from spaCy - \textit{en\_vectors\_web\_lg}, which contains 300-dimensional word vectors trained on Common Crawl with GloVe\cite{spacymodel}.
